### 2025-02-20 manuscript.Rmd
some more changes during revision

The ´reproducibleRchunks` package addresses the challenge of ensuring computational results in R Markdown documents are automatically testable for successful reproduction. 
This means verifying whether the same script with the same data yields identical results, even on different computers or at later times, with minimal disruption to users' workflows.
This article is intended for readers who use R for statistical analyses and ideally already had some experience with R Markdown. 
We aim to first offer a less technical tutorial outlining a typical use case for the package. 
Following this, we provide more detailed technical descriptions of the package's mechanics and various ways to customize its behavior.
We believe that the package can be used effectively without a deep understanding of the technical details. 
However, gaining insight into these details may enhance the user experience and help readers understand the challenges involved in ensuring reproducibility.

@@

On load, the package registers a new type of code chunk called `reproducibleR`.
Figure \@ref(fig:rstudioscreen1) shows a snippet from an R Markdown file, in which there is a `reproducibleR` chunk with the name _addition_. 
In the chunk, a new variable `my_sum` is declared and defined to be the sum of some variable `x` plus one. 
Variable `x` was declared in an earlier code chunk (not shown in the figure) and is thus not subject to the reproducibility test of this chunk.

@@

If the document is reproducedj and the computational result differs from the original result, this failure to reproduce is noted in the reproducibility report.
In a variation of the previous example, we assume someone changed the value of `x`, then variable `my_sum` would change and the a failure would be displayed (see Figure \@ref(fig:rstudioscreen4)).

@@

Ideally, the reproducibility metadata files are always kept together with the original Markdown file. For example, when providing Open Code on open platforms, the repository should contain both the Markdown files as well as the metadata files. Note that the metadata file format is particularly suitable to be used with version control systems [see @peikert2021reproducible].

@@

The JSON format has a clear, formal structure, which allows easy parsing of the information both for computers and humans. 
Without going into detail, for each object, its type (e.g., character, integer, list, or class), its value and its attributes are stored. The stored values themselves values  can in turn consist of collections of objects (e.g., lists of lists); in this way, arbitrarily complex R objects (such as an entire regression model) can be represented.
For privacy reasons and to save disk space, we actually do not store the raw data by default.
Instead, we store "fingerprints" of the data.
Digital fingerprints are fixed-length representations of an arbitrarily large digital object.
They are generated such that small changes in the input object will lead to different fingerprints and, thus, they are ideally suited to verify integrity of the original data.
The idea is very similar to a check sum digit in a credit card number, which can be computed from the remaining digits to check whether the credit card number is valid.
Fingerprints are realized via one-way hash functions, which can take an arbitrary large digital object and map it onto a fixed-size object, which is typically displayed as hexadecimal string.
As default, we chose the SHA256 algorithm, which allows us to map any content of a variable in R to a 256 bit-long fingerprint or "hash", which is usually displayed using 64 hexadecimal characters.
The following shows the SHA256-fingerprint of the content of variable `numbers` from the previous example:
\begin{verbatim}
```{r echo=FALSE, eval=TRUE, results="asis"}
content <- jsonlite::serializeJSON(digest::digest(numbers,algo = "sha256"),pretty=TRUE)
content <- gsub(pattern="\\n",replacement="  \n", x=content)
cat(content)
```
\end{verbatim}
And the following example is a SHA256 fingerprint of the first chapter of Genesis from the Old Testament:
\begin{verbatim}
```{r echo=FALSE, eval=TRUE, results="asis"}
library(logos)
genesis <- select_passage("Genesis",chapter = 1, language = "English", testament = "old")
content <- jsonlite::serializeJSON(digest::digest(genesis,algo = "sha256"),pretty=TRUE)
content <- gsub(pattern="\\n",replacement="  \n", x=content)
cat(content)
```
\end{verbatim}
Note that these exemplary fingerprints have the same length despite both variables having different variable types and sizes.
Due to their small size, fingerprints allow for an efficient comparison of identity of results while avoiding data security and privacy issues as they do not allow the original objects to be reproduced.

>>>>

The ´reproducibleRchunks` package addresses the challenge of ensuring computational results in R Markdown documents are automatically testable for successful reproduction. 
This means verifying whether the same script with the same data yields identical results, even on different computers or at later times, with minimal disruption to users' workflows.
This article is intended for readers who use R for statistical analyses and ideally already had some experience with R Markdown. 
We aim to first offer a less technical tutorial outlining a typical use case for the package. 
Following this, we provide more detailed technical background information about the package and various ways to customize its behavior.
We believe that the package can be used effectively without a deep understanding of the technical details. 
However, gaining insight into these details may enhance the user experience and help readers understand the challenges involved in ensuring reproducibility.

@@

On load, the package registers a new type of code chunk called `reproducibleR`.
Figure \@ref(fig:rstudioscreen1) shows a snippet from an R Markdown file, in which there is a `reproducibleR` chunk with the name _addition_. The entire tutorial file can be found in the R package in sub directory "inst/examples_simple_demo" or can be downloaded from \url{https://github.com/brandmaier/reproducibleRchunks/tree/main/inst/examples/simple_demo}.
In the chunk, a new variable `my_sum` is declared and defined to be the sum of some variable `x` plus one. 
Variable `x` was declared in an earlier code chunk (not shown in the figure) and is thus not subject to the reproducibility test of this chunk.

@@

If the document is reproduced and the computational result differs from the original result, this failure to reproduce is noted in the reproducibility report.
In a variation of the previous example, we assume someone changed the value of `x`, then variable `my_sum` would change and the a failure would be displayed (see Figure \@ref(fig:rstudioscreen4)).

@@

Ideally, the reproducibility metadata files are always kept together with the original Markdown file. For example, when providing Open Code on open platforms, the repository should contain both the Markdown files as well as the metadata files. Note that the metadata file format is particularly suitable to be used with version control systems [see @peikert2021reproducible]. If users intentionally modify R code later, e.g., because they want to fix an error, then the metadata files would need to be deleted and recreated. Version control can help users in documenting such changes.

@@

The JSON format has a clear, formal structure, which allows easy parsing of the information both for computers and humans. 
Without going into detail, for each object, its type (e.g., character, integer, list, or class), its value and its attributes are stored. The stored values themselves values  can in turn consist of collections of objects (e.g., lists of lists); in this way, arbitrarily complex R objects (such as an entire regression model) can be represented.
For privacy reasons and to save disk space, we actually do not store the raw data by default.
Instead, we store "fingerprints" of the data.
Digital fingerprints are fixed-length representations of an arbitrarily large digital object.
They are generated such that small changes in the input object will lead to different fingerprints and, thus, they are ideally suited to verify integrity of the original data.
The idea is very similar to a check sum digit in a credit card number, which can be computed from the remaining digits to check whether the credit card number is valid.
Fingerprints are realized via one-way hash functions, which can take an arbitrary large digital object and map it onto a fixed-size object, which is typically displayed as hexadecimal string.
As default, we chose the SHA256 algorithm, which allows us to map any content of a variable in R to a 256 bit-long fingerprint or "hash", which is usually displayed using 64 hexadecimal characters.
The following shows the SHA256-fingerprint of the content of variable `numbers` from the previous example:
\begin{verbatim}
```{r echo=FALSE, eval=TRUE, results="asis"}
#content <- jsonlite::serializeJSON(digest::digest(numbers,algo = "sha256"),pretty=TRUE)
content <- digest::digest(numbers,algo = "sha256")
content <- gsub(pattern="\\n",replacement="  \n", x=content)
cat(content)
```
\end{verbatim}
And the following example is a SHA256 fingerprint of the first chapter of Genesis from the Old Testament:
\begin{verbatim}
```{r echo=FALSE, eval=TRUE, results="asis"}
library(logos)
genesis <- select_passage("Genesis",chapter = 1, language = "English", testament = "old")
#content <- jsonlite::serializeJSON(digest::digest(genesis,algo = "sha256"),pretty=TRUE)
content <- digest::digest(genesis,algo = "sha256")
content <- gsub(pattern="\\n",replacement="  \n", x=content)
cat(content)
```
\end{verbatim}
Note that these exemplary fingerprints have the same length despite both variables having different variable types and sizes.
Due to their small size, fingerprints allow for an efficient comparison of identity of results while avoiding data security and privacy issues as they do not allow the original objects to be reproduced.
